---
title: "Nutrien_stats"
author: "mahafujul_hamid_ananda"
date: "2024-01-07"
output:
  word_document: default
  html_document: default
---

This is a Statistical Analysis of Fertilizer Sales of Nutrien from first Quarter of 2018 to third Quarter of 2023. The dataset as been extracted from Nutrien's website.

Loading the libraries

```{r}
library(openxlsx)
library(readxl)
library(tidyverse)
library(dplyr)
library(tidyr)
```


Now i will be loading the dataset from my directory to the Rstudio and select the Segment Summary (Q) sheet and Income Statement (Q) sheet to perform statistical analysis.


```{r}
# Load the workbook and select the relevant sheets
workbook <- loadWorkbook("dataset/nutrien.xlsx")
sheetNames <- getSheetNames("dataset/nutrien.xlsx")
sheetNames

segmentSummary <- read.xlsx(workbook, sheet = "Segment Summary (Q)")
incomeStatement <- read.xlsx(workbook, sheet = "Income Statement (Q)")

```


After selecting the sheets to work on, I am extracting the rows and columns for quarters, potash, nitrogen, phosphate, and net earnings.

```{r}
# Extract specific rows and columns for quarters, potash, nitrogen, phosphate, and net earnings
quarters <- segmentSummary %>%
  slice(1:1) %>% 
  select(Q1.18:Q3.23)

potash <- segmentSummary %>%
  slice(27:27) %>%
  select(Q1.18:Q3.23)
potash <- as.integer(potash)

nitrogen <- segmentSummary %>%
  slice(46:46) %>%
  select(Q1.18:Q3.23)
nitrogen <- as.integer(nitrogen)

phosphate <- segmentSummary %>%
  slice(65:65) %>%
  select(Q1.18:Q3.23)
phosphate <- as.integer(phosphate)

netEarning <- incomeStatement %>%
  slice(20:20) %>%
  select(Q1.18:Q3.23)
netEarning <- as.integer(netEarning)

```

Now i will combine the extracted data and prepare it for plotting and analysis by first transposing the data and then changing the data types of columns.

```{r}
# Combine extracted data into a single data frame
fertilizerQuarterlyData <- rbind(quarters, potash, nitrogen, phosphate, netEarning)
rownames(fertilizerQuarterlyData)[which(rownames(fertilizerQuarterlyData) == "1")] <- "Quarters"
rownames(fertilizerQuarterlyData)[which(rownames(fertilizerQuarterlyData) == "2")] <- "Potash"
rownames(fertilizerQuarterlyData)[which(rownames(fertilizerQuarterlyData) == "3")] <- "Nitrogen"
rownames(fertilizerQuarterlyData)[which(rownames(fertilizerQuarterlyData) == "4")] <- "Phosphate"
rownames(fertilizerQuarterlyData)[which(rownames(fertilizerQuarterlyData) == "5")] <- "Net_Earning"

# Prepare the data for plotting
fertilizerQuarterlyDataFlipped <- t(fertilizerQuarterlyData)
fertilizerQuarterlyDataFlipped <- as.data.frame(fertilizerQuarterlyDataFlipped)

# Convert Quarters to date format and rename columns
fertilizerQuarterlyDataFlipped$Quarters <- as.Date(fertilizerQuarterlyDataFlipped$Quarters, format="%m/%d/%Y")
fertilizerQuarterlyDataFlipped$Quarters <- format(fertilizerQuarterlyDataFlipped$Quarters, "%Y-%m")

# Convert data types
fertilizerQuarterlyDataFlipped$Potash <- as.integer(fertilizerQuarterlyDataFlipped$Potash)
fertilizerQuarterlyDataFlipped$Nitrogen <- as.integer(fertilizerQuarterlyDataFlipped$Nitrogen)
fertilizerQuarterlyDataFlipped$Phosphate <- as.integer(fertilizerQuarterlyDataFlipped$Phosphate)
fertilizerQuarterlyDataFlipped$Net_Earning <- as.integer(fertilizerQuarterlyDataFlipped$Net_Earning)

# Prepare data for plotting - pivot longer
fertilizerQuarterlyDataFlippedMerged <- pivot_longer(fertilizerQuarterlyDataFlipped, cols = c("Potash","Nitrogen","Phosphate"), names_to= "Fertilizer", values_to =  "Sales" )


```

After preparing the data, i am plotting a bar chart for the fertilizer sales throughout the quarters

```{r echo=FALSE,message=FALSE, fig.height=8, fig.width=14, fig.show='hold'}
# Create a bar chart
barChart <- ggplot()
barChart <- barChart + geom_col(data = fertilizerQuarterlyDataFlippedMerged, aes(x = Quarters, y = Sales, fill= Fertilizer ), position =  "dodge" ) +
  labs(x = "Quarters", y = "Net Fertilizer Sales (millions)", title = "Net Fertilizer Sales Over Quarters") +
  theme(axis.text = element_text(size = 7)) + # Adjust size as needed
  scale_y_continuous(
    breaks = seq(0, 3000, by = 200),  # Custom breaks at 0, 500, 1000, 1500, 2000, 2500, 3000
  )
barChart
```


Now, I am creating a line graph to analyze the revenue trend

```{r echo=FALSE,message=FALSE, fig.height=8, fig.width=14, fig.show='hold'}
# Create a line graph with highlighted points
lineGraph <- ggplot()
lineGraph <- ggplot(fertilizerQuarterlyDataFlippedMerged, aes(x = Quarters, y = Net_Earning, group=1)) +
  geom_line() +  # Specify the type of plot (in this case, a line plot)
  geom_point() + # Add points for each data point
  labs(x = "Quarters", y = "Net Earnings (millions)", title = "Net Earnings Over Quarters") +  # Label axes and title
  theme(axis.text = element_text(size = 7)) + # Adjust size as needed
  theme_minimal() + # Set a minimal theme (you can choose a different theme based on your preference)
  scale_y_continuous(
    breaks = seq(-1200, 4000, by = 200),  # Custom breaks at 0, 500, 1000, 1500, 2000, 2500, 3000
  )

# Find max and min points
max_point <- fertilizerQuarterlyDataFlippedMerged[which.max(fertilizerQuarterlyDataFlippedMerged$Net_Earning), ]
min_point <- fertilizerQuarterlyDataFlippedMerged[which.min(fertilizerQuarterlyDataFlippedMerged$Net_Earning), ]

# Extract quarters for max and min points
max_quarter <- max_point$Quarters
min_quarter <- min_point$Quarters

# Highlight max and min points on the line graph
lineGraph +
  geom_point(data = rbind(max_point, min_point), aes(x = Quarters, y = Net_Earning, color = c("Max", "Min")), size = 3) +
  scale_color_manual(values = c(Max = "green", Min = "red")) +
  geom_text(data = rbind(max_point, min_point), aes(x = Quarters, y = Net_Earning, label = Net_Earning),
            vjust = ifelse(max_point$Net_Earning > min_point$Net_Earning, -0.5, 1.5), hjust = ifelse(max_point$Net_Earning > min_point$Net_Earning, 1.5, -0.5), size = 3, color = c("green", "red")) +
  guides(color = FALSE)  # Hide the legend for the annotations

```

After plotting the graphs, i will now do some statistical analysis, such as the mean, maximum and minimum sales of fertilizers as well as the mean, maximum and minimum earnings,on the data i gathered from Nutrien.

```{r}

# Calculate and display statistical information for Net Earnings
averageNetEarnings <- mean(netEarning)
cat("The mean Net Earning is", averageNetEarnings, "millions", "\n\n")

maxNetEarning <- max(netEarning)
maxEarningColumnIndex <- which(fertilizerQuarterlyData[5, ] == maxNetEarning)
maxSalesQuarterEarning <- fertilizerQuarterlyData[1, maxEarningColumnIndex]
cat("Maximun Net Earning was" , maxNetEarning , " millions, and the Earning Quarter is" , maxSalesQuarterEarning, "\n\n" )

minNetEarning <- min(netEarning)
minEarningColumnIndex <- which(fertilizerQuarterlyData[5, ] == minNetEarning)
minSalesQuarterEarning <- fertilizerQuarterlyData[1, minEarningColumnIndex]
cat("Minimun Net Earning was" , minNetEarning , " millions, and the Earning Quarter is" , minSalesQuarterEarning, "\n\n" )

# Calculate and display statistical information for Potash sales
averagePotashSales <- mean(potash)
cat("The mean Potash sales is", averagePotashSales, "millions", "\n\n")

maxPotashSale <- max(potash)
maxPotashColumnIndex <- which(fertilizerQuarterlyData[2, ] == maxPotashSale)
maxSalesQuarterPotash <- fertilizerQuarterlyData[1, maxPotashColumnIndex]
cat("Maximun Potash sale was" , maxPotashSale , " millions, and the Sales Quarter is" , maxSalesQuarterPotash, "\n\n" )

minPotashSale <- min(potash)
minPotashColumnIndex <- which(fertilizerQuarterlyData[2, ] == minPotashSale)
minSalesQuarterPotash <- fertilizerQuarterlyData[1, minPotashColumnIndex]
cat("Minimun Potash sale was" , minPotashSale , " millions, and the Sales Quarter is" , minSalesQuarterPotash, "\n\n")

# Calculate and display statistical information for Nitrogen sales
averageNitrogenSales <- mean(nitrogen)
cat("The mean Nitrogen sales is", averageNitrogenSales, "millions", "\n\n")

maxNitrogenSale <- max(nitrogen)
maxNitrogenColumnIndex <- which(fertilizerQuarterlyData[3, ] == maxNitrogenSale)
salesQuarterNitrogen <- fertilizerQuarterlyData[1, maxNitrogenColumnIndex]
cat("Maximun Nitrogen sale was" , maxNitrogenSale , " millions, and the Sales Quarter is" , salesQuarterNitrogen, "\n\n" )

minNitrogenSale <- min(nitrogen)
minmaxNitrogenColumnIndex <- which(fertilizerQuarterlyData[3, ] == minNitrogenSale)
minSalesQuarterNitrogen <- fertilizerQuarterlyData[1, minmaxNitrogenColumnIndex]

cat("Minimun Nitrogen sale was" , minNitrogenSale , " millions, and the Sales Quarter is" , minSalesQuarterNitrogen, "\n\n" )

# Calculate and display statistical information for Phosphate sales
averagePhosphateSales <- mean(phosphate)

cat("The mean Phosphate sales is", averagePhosphateSales, "millions", "\n\n")

maxPhosphateSale <- max(phosphate)
maxPhosphateColumnIndex <- which(fertilizerQuarterlyData[4, ] == maxPhosphateSale)
salesQuarterPhosphate <- fertilizerQuarterlyData[1, maxPhosphateColumnIndex]

cat("Maximun Phosphate sale was" , maxPhosphateSale , " millions, and the Sales Quarter is" , salesQuarterPhosphate, "\n\n" )

minPhosphateSale <- min(phosphate)
minmaxPhosphateColumnIndex <- which(fertilizerQuarterlyData[4, ] == minPhosphateSale)
minSalesQuarterPhosphate <- fertilizerQuarterlyData[1, minmaxPhosphateColumnIndex]

cat("Minimun Phosphate sale was" , minPhosphateSale , " millions, and the Sales Quarter is" , minSalesQuarterPhosphate, "\n\n" )

```

From the data analysis we learned that on second quarter of 2022 we had the highest sales of Potash and Nitrogen and it was also the quarter where we had maximum Net Earnings!

